---
title: "K-Means for Clustering"
author: "Dutt Thakkar"
date: "2023-03-19"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#importing data
```{r}
library(readr)
Pharm = read.csv("/Users/duttthakkar/Desktop/Pharm.csv")
df= Pharm
```

#viewing the summary of the dataset
```{r}
summary(df)
```

#attaching required libraries
```{r}
library(tinytex)
library(tidyverse)
library(ISLR)
library(flexclust)
library(FactoMineR)
library(ggcorrplot)
library(ggplot2)
library(factoextra)
```

## Question A: Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.

#subsetting the data
```{r}
names(df)
```
```{r}
dataset=df[,c(1,3:11)]
row.names(dataset)=dataset[,1]
dataset=dataset[,-1]
head(dataset)
```
#normalizing the dataset
```{r}
dataset2 = scale(dataset)
distance=get_dist(dataset)
fviz_dist(distance)
```
#using euclidean distance formula which is given by: d=√((x_2-x)1)^2+(y_2-y_1 )^2 )
```{r}
Corr=cor(dataset2)
ggcorrplot(Corr,outline.color = "black",lab = TRUE,hc.order = TRUE,type = "full")
```
#installing factoextra and cluster to plot elbow chart and silhouette chart
```{r}
library(cluster)
library(factoextra)
```
#Finding the number of clusters using elbow chart 
```{r}
set.seed(100)
fviz_nbclust(dataset2, kmeans, method = "wss")
```
#after looking at the elbow chart, it shows that the optimal number of clusters is 2 or 7
#Finding the number of clusters using silhouette method
```{r}
fviz_nbclust(dataset2,kmeans,method = "silhouette")
```
#after looking at the elbow chart, it shows that the optimal number of clusters is 5. Therefore, we will try and find an optimal value between 2 and 7 per the results gathered from elbow and silhouette method respectively

```{r}
k2<-kmeans(dataset2,centers =2,nstart=25)
k3<-kmeans(dataset2,centers =3,nstart=25)
k4<-kmeans(dataset2,centers =4,nstart=25)
k5<-kmeans(dataset2,centers =5,nstart=25)
k6<-kmeans(dataset2,centers =6,nstart=25)
k7<-kmeans(dataset2,centers =7,nstart=25)
p1<-fviz_cluster(k2,geom = "point", data=dataset2)+ggtitle("k=2")
p2<-fviz_cluster(k3,geom = "point", data=dataset2)+ggtitle("k=3")
p3<-fviz_cluster(k4,geom = "point", data=dataset2)+ggtitle("k=4")
p4<-fviz_cluster(k5,geom = "point", data=dataset2)+ggtitle("k=5")
p5<-fviz_cluster(k6,geom = "point", data=dataset2)+ggtitle("k=6")
p6<-fviz_cluster(k7,geom = "point", data=dataset2)+ggtitle("k=7")
```

#attaching library gridExtra to combine the clusters
```{r}
library(gridExtra)
grid.arrange(p1,p2,p3,p4,p5,p6)
```
##Question B: Interpret the clusters with respect to the numerical variables used in forming the clusters. 
# after reviewing the clusters, K = 5 seems appropriate as per the grouping. 

#using K=5 for the analysis
```{r}
k5=kmeans(dataset2, centers = 5, nstart = 25)
k5$size
k5$cluster
k5$centers
k5$withinss
fviz_cluster(k5, data = dataset2)
```
#Interpretation of the clusters
#The entire data is divided into 5 different clusters: Cluster #5 have the 4 companies and their net profit margin is high as well as their asset turnover making them a credible institutions. On the other hand, cluster #1 has 3 companies and their net profit margin is -1.32 as they are more dependent on levraging (1.36) and less on equities (-0.61)

#plotting graphs of data grouped in clusters
```{r}
Centroid <- data.frame(k5$centers) %>% rowid_to_column() %>% gather('Columns', 'Centers', -1)
print(Centroid)
ggplot(Centroid, aes(x = Columns, y = Centers, color = as.factor(rowid))) + geom_line(aes(group = as.factor(rowid))) + geom_point()
```
#Question C: Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? 

#considering the last three variables; median_recommendation, location, and exchange
```{r}
pattern <- df%>% select(c(12,13,14)) %>% mutate(Cluster = k5$cluster)
print(pattern)
```
#identifying if there are any trends 
```{r}
Median_Recommenation <- ggplot(pattern, mapping = aes(factor(Cluster), fill=Median_Recommendation)) + geom_bar(position = 'dodge') + labs(x='Clusters', y='Frequence')
Location <- ggplot(pattern, mapping = aes(factor(Cluster), fill=Location)) + geom_bar(position = 'dodge') + labs(x='Clusters', y='Frequence')
Exchange <- ggplot(pattern, mapping = aes(factor(Cluster), fill=Exchange)) + geom_bar(position = 'dodge') + labs(x='Clusters', y='Frequence')
grid.arrange(Median_Recommenation,Location,Exchange)
```

```{r}
#Cluster1 has majority of the companies from the US listed equally in AMEX, NASDAQ, and NYSE. This segment contains low-risk companies as their holding rate is higher than the buying rates.
#Cluster2 has majority of the companies the US followed by UK and Switzerland. All the companies are listed in the NYSE. These companies are moderately low-risk companies as their holding rate is still higher but also shows adequate selling.
#Cluster3 has companies from Canada and US, listed in NYSE. These companies demonstrates some growth potential as equal number of holding and buying rates.
#Cluster 4 has companies from France, Germany and US investing at NYSE.These companies shows the most risky activities as they have equal buying and selling rates. On the contrary, this shows that as they take the risk, they have higher potential of growth.
#Cluster 5 has companies US and UK again listed in NYSE. These companies practice the safest among all the clusters. Their holding and buying rates are equal but slightly higher that cluster#3. These are the most profitable companies. 
```

#Question D: Provide an appropriate name for each cluster using any or all of the variables in the dataset.

```{r}
#Cluster 1: Low-risk companies (well-ordered)
#Cluster 2: Growing companies
#Cluster 3: high-risk companies
#Cluster 4: risky-companies
#Cluster 5: Stable companies
#These titles has been given after comparing the five different clusters and their K-center values (Market_Cap, Beta, PE_Ratio, ROE, ROA, Asset_Turnover, Leverage, Rev_Growth, Net_Profit_Margin ). 
```

