---
title: "Assignment"
output: word_document
date: "2023-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Assignment 1 - Introduction to R
# Source of dataset - Kaggle.com 
# https://www.kaggle.com/datasets/ritesh2000/trending-tv-shows-on-netflix
# Title of the dataset: Trending TV shows on Netflix
#Data Content:
#Title: Name of the TV show
#Year: Year in which the show was released
#Rating: Ratings given by Netflix
#IMDB_Ratings: Ratings by IMDB
#Netflix: If the show is currently streaming on Netflix or not
```{r}
library(readxl)
Netflix_data <- read_excel("Netflix_data.xls")
View(Netflix_data)
```
#Question 1:
#Print out descriptive statistics for quantitative and qualitative variables
```{r}
summary(Netflix_data)
table(Netflix_data$Year)
table(Netflix_data$Titles)
library(psych)
describe(Netflix_data$Year)
```
#Question2:
#Transforming categorical variable into numeric variable
```{r}
Netflix_data$Titles=as.numeric(Netflix_data$Titles)
as.numeric(Netflix_data$Titles)
```
#Categorical variable was converted to N/A as the data was non-numeric.
#Tranforming numeric variable into categorical varaible
```{r}
attach(Netflix_data)
names(Netflix_data)
Year[1:10]
CatYear <- cut(Year, breaks = c(1990,1995,2000,2005,2010,2015,2020), labels = c("A", "B","C","D","E","F"))
CatYear[1:5]
```
#Tranforming categorical variable into factor (numeric) variable
```{r}
Titles=c("Breaking Bad","Game of Thrones","Breaking Bad","Game of Thrones","Rick and Morty","Dark","Stranger Things","Avatar: The Last Airbender","Sherlock","Chernobyl","Friends","The Office","Better Call Saul","The Wire","Supernatural","Fargo","Black Mirror","Band of Brothers","Attack on Titan","House","Money Heist","Peaky Blinders","Mr. Robot","Vikings","The Sopranos","The Simpsons","The Mandalorian","Lost","Westworld","The Umbrella Academy","Narcos","True Detective","Marvel's Daredevil","Brooklyn Nine-Nine","The Walking Dead","Parks and Recreation","Mindhunter","South Park","Suits","Hannibal","Dexter","Community","Fringe","Modern Family","The Big Bang Theory","Firefly","Mad Men","Fullmetal Alchemist: Brotherhood","Ozark","The Witcher","Lucifer","It's Always Sunny in Philadelphia")
Titles_factor<-factor(Titles)
Titles_factor
```
#Question 3:
#Plot at least one quantitative variable, and one scatter plot
#Histogram
```{r}
variable=Netflix_data$Year
hist(Netflix_data$Year, main = "Year of the Show", col = ("blue"))
```
#Scatter Plot
```{r}
plot(Netflix_data$Year,Netflix_data$IMDB_Rating, main = "Ratings per year", col="green")
abline(lm(IMDB_Rating~Year), col=("red"))
```
