---
title: "ML Final Project Extra Credit"
author: "Dutt Thakkar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# For Regression model, using different data subset for evaluating the regression results. Selecting those variables that acutally affects the analysis

#loading reqiured pacakges
```{r}
library(caret)
library(tidyverse)
library(tidyr)
library(dplyr)
library(factoextra)
library(cluster)
```

#importing dataset and viewing summary
```{r}
Data_set <- read.csv("/Users/duttthakkar/Desktop/fuel_receipts_costs_eia923(1).csv")
summary(Data_set)
```
#data cleaning

#selecting attributes
```{r}
fuel_data<-Data_set[,c(11,16,17,18,20)]
summary(fuel_data)
str(fuel_data)
```
#checking for Na's
```{r}
colMeans(is.na(fuel_data))
```

#Data imputing
```{r}
fuel_data$fuel_cost_per_mmbtu[is.na(fuel_data$fuel_cost_per_mmbtu)] <- mean(fuel_data$fuel_cost_per_mmbtu, na.rm = TRUE)
colMeans(is.na(fuel_data))
```
#all Na's have been imputed using the mean 

#Data partition
```{r}
library(caTools)
set.seed(2299)
# Sample about 2% of data
sample_size <- round(0.02 * nrow(fuel_data))
sample_indices <- sample(nrow(fuel_data), sample_size, replace = FALSE)

# Split sampled data into training and test sets
train_data <- fuel_data[sample_indices[1:round(0.75*sample_size)], ]
test_data <- fuel_data[sample_indices[(round(0.75*sample_size) + 1):sample_size], ]

nrow(train_data)
nrow(test_data)
```

#normalization of the data
```{r}
cluster_data <- train_data %>% select( 'ash_content_pct', 'sulfur_content_pct','fuel_mmbtu_per_unit','fuel_cost_per_mmbtu')

cluster_train <- preProcess(cluster_data, method = "range")
cluster_predict <- predict(cluster_train, cluster_data)

summary(cluster_predict)
```



#Applying K-means Algorithm
```{r}
KMean_ <- kmeans(cluster_predict, centers = 2, nstart = 30) 
```

#centers
```{r}
KMean_$centers
```


#The final cluster
```{r}
fcluster<- KMean_$cluster
f_cluster<- cbind(train_data, fcluster)
f_cluster$fcluster<-as.factor(f_cluster$fcluster)
head(f_cluster)
```
#We find the mean of all the quantitative variables
```{r}
f_cluster%>%group_by(fcluster)%>%
  summarize(
    fuel_mmbtu_per_unit=mean(fuel_mmbtu_per_unit),
            fuel_cost_per_mmbtu=mean(fuel_cost_per_mmbtu),
            sulfur_content=mean(sulfur_content_pct),
    ash_content=mean(ash_content_pct))
```


#Use multiple-linear regression to determine the best set of variables to predict fuel_cost_per_mmbtu

#training data
```{r}
reg_df<- f_cluster
fuel<-reg_df[,-c(1)]
fuel_ML<- preProcess(fuel, method = "range")
fuel_predict <- predict(fuel_ML, fuel)
head(fuel_predict)
```
#performing multiple linear regression model on training data
```{r}
k<-fuel_predict$fuel_cost_per_mmbtu
D1<- fuel_predict$fuel_mmbtu_per_unit 
D2<- fuel_predict$sulfur_content_pct
D3<- fuel_predict$ash_content_pct
model_check <- lm(fuel_cost_per_mmbtu~.,data=fuel_predict)
summary(model_check)

```
#Use the anova analysis
```{r}
anova(model_check)
```

#Test data
```{r}
Check_df<- test_data
fuel<-Check_df[,-c(1)]
fuel_chk<- preProcess(fuel, method = "range")
fuel_check <- predict(fuel_chk, fuel)
head(fuel_check)
```

#performing multiple linear regression model on test data
```{r}
M<-fuel_check$fuel_cost_per_mmbtu

T1<- fuel_predict$fuel_mmbtu_per_unit 
T2<- fuel_predict$sulfur_content_pct
T3<- fuel_predict$ash_content_pct
model_check1 <- lm(fuel_cost_per_mmbtu~.,data=fuel_check)
summary(model_check1)
```
#Use the anova analysis to predict the model
```{r}
anova(model_check1)
```
# Conclusion

#It appears that the predictors in your model have varying levels of significance. The intercept and fuel_mmbtu_per_unit have statistically significant coefficients, while sulfur_content_pct and ash_content_pct do not appear to have a statistically significant effect on fuel_cost_per_mmbtu. Additionally, the fcluster2 variable also has a statistically significant effect on fuel_cost_per_mmbtu. In the test data, only the fuel_mmbtu_per_unit predictor appears to be statistically significant, while sulfur_content_pct and ash_content_pct do not appear to have a significant effect on fuel_cost_per_mmbtu. Overall, it seems that fuel_mmbtu_per_unit is the most important predictor in your model for predicting fuel_cost_per_mmbtu, with the other predictors having limited impact.